Pre-Main-Sequence M-Dwarfs With Complex Quasiperiodic Variability From Four Years of TESS

### 2023-06-14 paper/ms.tex
update ms

\title{Complex Quasiperiodic Pre-Main-Sequence M-Dwarfs from Four Years of TESS}

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are pre-main-sequence
  M-dwarfs with almost-periodic optical modulation that includes both
  sharp and broad troughs.  The troughs are usually superposed over
  smooth, spot-like modulation.  These objects do not have primordial
  gas-rich disks, based on a lack of observed infrared excesses.
  Nonetheless, the currently preferred explanations for their optical
  variability invoke long-lived material in orbit at the corotation
  radius, for instance transiting dust clumps or stellar prominences.
  Here, we report new CQVs discovered in TESS data collected between
  July~2018 and Sep~2022.  Our search used two methods, the first
  based on counting dips in phase-folded light curves, and the second
  based on counting peaks in the Fourier domain.  From our analysis of
  \nstarssearched\ M-dwarfs with $T$$<$16 and $d$$<$150\,pc, we find
  \ncpvsfound\ gold-standard CQVs.  Most of these discoveries are new,
  and they include the brightest ($T$$\approx$10) and closest
  ($d$$\approx$20\,pc) examples of this class of object known.  A few
  objects are ``outliers among outliers''; for instance LP 12-502
  shows four to eight local minima per cycle, distributed over a ``dip
  complex'' with a total duration that is fixed over more than one
  thousand cycles, but with a detailed shape that exhibited changes
  over less than one cycle.  More generally, we find that none of the
  CQVs maintain a fixed light curve shape over timescales of more than
  a few hundred cycles -- hence, {\it quasi}periodic.  Although the
  physical explanation behind these objects remains enigmatic, our
  sample may help clarify their origin.  Connections to star, disk,
  and exoplanet evolution are emphasized.
\end{abstract}

@@

@@

We chose to analyze only the ``short'' 120-second cadence data
acquired by TESS between July~2018 and Sep~2022 (Sectors 1-55).
Specifically, we used the 120-second cadence light curve products
produced by the mission's Science Processing and Operations Center at
NASA Ames \citep{2016SPIE.9913E..3EJ}.  While the TESS data products
also include full frame images with cadences of 200, 600, and 1800
seconds, we limited our scope in this work for the sake of simplicity
in data handling.  In exchange, we lose out in both completeness and
homogeneity of the selection function.  While TESS cumulatively
observed $\approx$90\% of the sky for at least one lunar month between
July~2018 and Sep~2022, the 120-second cadence data were acquired for
only a pre-selected set of stars over Sectors 1-26, and then a
guest-investigator driven set of stars over Sectors 27-55
\citep{2021PASP..133i5002F}.

>>>>

\title{Pre-Main-Sequence M-Dwarfs With Complex Quasiperiodic Variability From Four Years of TESS}

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are pre-main-sequence
  M-dwarfs with almost-periodic optical modulation that includes both
  sharp and broad troughs.  The troughs are usually superposed over
  smooth, spot-like modulation.  These objects are typically
  10-100\,million years old, and they do not seem to have primordial gas-rich
  disks.  Nonetheless, the preferred explanations for their optical
  variability invoke long-lived material in orbit at the corotation
  radius, for instance transiting dust clumps or stellar prominences.
  Here, we report new CQVs discovered in TESS data collected between
  July~2018 and Sep~2022.  Our search used two methods, the first
  based on counting dips in phase-folded light curves, and the second
  based on counting peaks in the Fourier domain.  From our analysis of
  \nstarssearched\ M-dwarfs with $T$$<$16 and $d$$<$150\,pc, we find
  \ncpvsfound\ gold-standard CQVs.  Most of these discoveries are new,
  and they include the brightest ($T$$\approx$10) and closest
  ($d$$\approx$20\,pc) examples of this class of object known.  A few
  objects are ``outliers among outliers''; for instance LP 12-502
  shows four to eight local minima per cycle distributed over a ``dip
  complex'' with a total duration that is nearly fixed over more than
  1000 cycles.  Nonetheless, LP-502 at times shows drastic shape
  changes over less than one cycle.  More generally, we find that none
  of the CQVs maintain a fixed light curve shape over timescales of
  more than a few hundred cycles -- hence, {\it quasi}periodic.
  Although the physical explanation behind these objects remains
  enigmatic, our sample may help clarify their origin.  Connections to
  star, disk, and exoplanet evolution are emphasized.
\end{abstract}

@@

% NOTE: could write this more from a stellar evolution angle.
% M-dwarfs contract.  They spin up.  Some reach rapid rotation rates,
% and exhibit strong magnetic fields.

@@

We analyzed the ``short'' 120-second cadence data
acquired by TESS between July~2018 and Sep~2022 (Sectors 1-55).
Specifically, we used the 120-second cadence light curve products
produced by the mission's Science Processing and Operations Center at
NASA Ames \citep{2016SPIE.9913E..3EJ}.  While the TESS data products
also include full frame images with cadences of 200, 600, and 1800
seconds, we limited our scope in this work for the sake of simplicity
in data handling.  In exchange, we lose out in both completeness and
homogeneity of the selection function.  While TESS cumulatively
observed $\approx$90\% of the sky for at least one lunar month between
July~2018 and Sep~2022, the 120-second cadence data were acquired for
only a pre-selected set of stars over Sectors 1-26, and then a
guest-investigator driven set of stars over Sectors 27-55
\citep{2021PASP..133i5002F}.
